|

### 2022-10-04 paper_comparison/paper_comparison/paper_comparison.Rmd

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. As shown in Figure \ref{fig:commondepartures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns. There are various types of departures from a good residual plot. Two most commonly checked departures - non-linearity, heterskedasticity and non-normality will be discussed in this paper.

Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation. Any non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity. An example residual plot containing visual pattern of non-linearity is given at the top-right of Figure  \ref{fig:commondepartures}. One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.

@@

```{=tex}
\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \int_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)}dx,
\end{equation}
```

@@

As discussed in @majumder_validation_2013, individual's skill will affect the number of observers who identify the actual data plot from the lineup. Thus, the power of a visual test depends on the subject-specific abilities. Previously, it was addressed by modelling the probability of a subject $i$ correctly picking the actual data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013]. However, knowing this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject, and hence, the p-value can't be computed.

(1) Instead, we modelled the probability of a lineup being rejected directly using a logistic regression with the natural logarithm of the effect size as the only regressor. The same model was applied to different subsets of the survey data for making comparison of the power.

(2) Instead, we modelled the plot selection $c_i$ for the actual data plot $i$.

<!-- -   Very briefly explain what a lineup test is -->

<!-- -   Lineup (maybe using your previous residual plot for nonlinearity) -->

<!-- -   $p$-value calulcation, briefly, including how multiple selections are handled -->

<!-- -   power calculation, briefly, including how multiple selections are handled -->

```{=html}
<!-- ![Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009]. Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null. For regression diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives. \label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
-->
```

@@

<!-- The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous. The example residual plots are given in Figure \ref{fig:different-sigma}. -->

```{r different-sigma, eval = FALSE, fig.cap="Example residual plots of the linear regression model used in experiment I. Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)

Four different distribution were used to generate $X_{raw}$ as shown in Figure \ref{fig:different-dist}. The uniform and the normal distribution are symmetric and commonly assumed in statistical models. The adjusted log-normal distribution provides skewed density. And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.

@@

### Treatments

To control the strength of the signal, two parameters were used including $\sigma$ and $b$ for controlling the standard deviation of the error term in experiment I and II respectively.

<!-- > Experiment I -->

<!-- -   Treatments for nonlinearity experiment (plots summarising levels) -->

<!-- It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$. For this lineup, x out of k observers picked the actual data plot. Each lineup is evaluated by 5 different subjects to provide reasonable estimates of the p-value. -->

<!-- > Experiment II -->

<!-- -   Treatments for heteroskedasticity experiment (plots summarising levels) -->

<!-- Experiment data were simulated using three different shapes ($a$ = -1, 0, 1), five different values of $b$ (b = 0.25, 1, 4, 16, 64), three different sample sizes (n = 50, 100, 300) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$. A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}. -->

<!-- The range of $b$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous. -->

### Assigning subjects

<!-- Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups. For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis. The 19 null datasets were generated by the residual rotation technique, and plotted in the same way. The lineup consisted of 20 residual plots with one randomly placed actual data plot. Figure \ref{fig:example-lineup} is an example of one of these lineups. It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $X_{raw} \sim N(0.0.3^2)$. The actual data plot location was four. All five subjects correctly identified the actual data plot for this lineup. -->

<!-- In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value. Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the design of the experiment I. -->

@@

software/technical

@@

@@

```{r}
vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  rename(p_value_dirichlet = p_value_select_all_plots) %>%
  ggplot() +
  geom_point(aes(p_value_uniform, p_value_dirichlet)) +
  geom_abline() +
  scale_x_log10() +
  scale_y_log10()
```

```{r}
vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                pivot_longer(F_p_value:SW_p_value), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  facet_wrap(~x_dist) +
  ggtitle("poly")
```

```{r}
vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(reject_uniform = as.numeric(p_value_uniform <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value_dirichlet"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value_dirichlet"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(aes(log_effect_size, reject_uniform, col = "VI_p_value_uniform"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_uniform, times = 500), 
              aes(log_effect_size, reject_uniform, group = boot_id, col = "VI_p_value_uniform"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  facet_wrap(~x_dist) +
  ggtitle("poly")
```

@@

```{r}
vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(reject_uniform = as.numeric(p_value_uniform <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value_dirichlet"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value_dirichlet"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(aes(log_effect_size, reject_uniform, col = "VI_p_value_uniform"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_uniform, times = 500), 
              aes(log_effect_size, reject_uniform, group = boot_id, col = "VI_p_value_uniform"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  facet_wrap(~x_dist) +
  ggtitle("heter")
```

```{r}
p_value_cache_env <- new.env()

if (file.exists(here::here("data/p_value_only_5_eval.rds"))) {
  p_value_only_5_eval <- readRDS(here::here("data/p_value_only_5_eval.rds"))
} else {
  p_value_only_5_eval <- vi_survey_processed %>%
  filter(x_dist == "uniform") %>%
  filter((is.na(b) & e_sigma != 0.125 ) | (b != 0 & b < 128)) %>%
  filter(exp != 5) %>%
  group_by(unique_lineup_id) %>%
  summarise(p_value_only_5_eval = calc_p_value(n_detect = sum(detect), 
                                               n_eval = n(), 
                                               n_sel = num_selection, 
                                               dist = "dirichlet",
                                               alpha = numeric_alpha[1],
                                               cache_env = p_value_cache_env))

  saveRDS(p_value_only_5_eval, here::here("data/p_value_only_5_eval.rds"))
}

vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(x_dist == "uniform") %>%
  filter(b != 0) %>%
  filter(b < 128) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  left_join(p_value_only_5_eval) %>%
  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  ggtitle("heter")

p1 <- vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(x_dist == "uniform") %>%
  filter(b != 0) %>%
  filter(b < 128) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  left_join(p_value_only_5_eval) %>%
  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03)

p2 <- vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(x_dist == "uniform") %>%
  filter(b != 0) %>%
  filter(b < 128) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  left_join(p_value_only_5_eval) %>%
  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03)

patchwork::wrap_plots(p1, p2)
```

vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(x_dist == "uniform") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  left_join(p_value_only_5_eval) %>%
  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  ggtitle("poly")

p1 <- vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(x_dist == "uniform") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  left_join(p_value_only_5_eval) %>%
  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03)

p2 <- vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(x_dist == "uniform") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  left_join(p_value_only_5_eval) %>%
  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03)

patchwork::wrap_plots(p1, p2)

```{r eval = FALSE}
vi_survey_processed %>%
  filter(exp == 3) %>%
  filter(e_sigma != 0.125) %>%
  mutate()

vi_survey_processed %>%
  filter(exp == 3) %>%
  filter(e_sigma != 0.125) %>%
  mutate(set = factor(set), lineup_id = factor(lineup_id)) %>%
  {lme4::glmer(weighted_detect ~ effect_size + ( 0 + effect_size | set) + ( 0 + effect_size | lineup_id), family = binomial ,data = .)} %>%
  predict(type = "response")

  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE)

>>>>

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. As shown in Figure \ref{fig:residual-plot-common-departures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns. There are various types of departures from a good residual plot.

Non-linearity, heterskedasticity and non-normality are perhaps the three mostly checked departures.
Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation. Any non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity. An example residual plot containing visual pattern of non-linearity is given at the top-right of Figure  \ref{fig:commondepartures}. One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.

@@

<!-- TODO:Clarify this is only for integer c_i -->

```{=tex}
\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \sum_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},
\end{equation}
```

@@

As discussed in @majumder_validation_2013, individual's skill will affect the number of observers who identify the actual data plot from the lineup. Thus, the power of a visual test depends on the subject-specific abilities. Previously, it was addressed by modelling the probability of a subject $i$ correctly picking the actual data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013]. However, having this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject for p-value calculation.

Instead, we directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor formulated as:

```{=tex}
\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
\end{equation}
```
where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.

Effect $E$ is derived from the Kullback-Leibler divergence (see [appendix ref here]) formulated as:

```{=tex}
\begin{equation} \label{eq:effect-size-ex1}
E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
\end{equation}
```
where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.

To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors. This includes [expansion].

@@

Four different distribution were used to generate $X_{raw}$ as shown in Figure \ref{fig:different-dist}. The uniform and the normal distribution are symmetric and commonly assumed in statistical models. The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.

@@

### Controlling the strength of the signal

As summarised by Table \ref{tab:model-parameter-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous. Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ were used in experiment I and II respectively. Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters.

Three different sample sizes were used (n = 50, 100, 300) in all three experiments. It can be observed from Figure \ref{fig:different-n} that with fewer data points drawn in a residual plot, the visual pattern is more difficult to be detected.

```{r different-n, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
ori_dat <- poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(300)

VI_MODEL$plot(ori_dat) + 
  ggtitle(quote(n~"="~300)) + 
  theme_light() -> p1

VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:50,], x, e))) +
  ggtitle(quote(n~"="~50)) + 
  theme_light() -> p2

VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:100,], x, e))) +
  ggtitle(quote(n~"="~100)) + 
  theme_light() -> p3

patchwork::wrap_plots(p2, p3, p1, ncol = 2)
```

```{r different-sigma, fig.cap="Example residual plots of the linear regression model used in experiment I. Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)

```{r different-b, fig.cap="Example residual plots of the linear regression model used in experiment II. Five different values of $b$ are included in the experiment to control the strength of the signal."}
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

ori_dat <- heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 0.25)$gen(300)

VI_MODEL$plot(ori_dat) +
  ggtitle(quote("b = 0.25")) +
  theme_light() -> p1

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 1)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 1")) +
  theme_light() -> p2

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 4)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 4")) +
  theme_light() -> p3

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 16)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 16")) +
  theme_light() -> p4

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 64)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 64")) +
  theme_light() -> p5

patchwork::wrap_plots(p1, p2, p3, p4, p5, ncol = 3)
```

### Subject allocation

Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 1152 different lineups. In addition, each lineup is designed to be evaluated by five different subjects and lineups with uniform distribution is designed to be evaluated by 11 subjects to provide reasonable estimates of the visual p-value.

Thus, $443$ subjects were recruited to satisfy the design of the experiment. 
<!-- TODO:subject allocation algorithm -->

@@

<!-- TODO:pipeline description -->

# Results

## Data overview

Subjects recruited from Prolific received a fixed payment for participating in the experiment. However, some subjects will try to maximize their earnings for minimum effort. During the review of submissions, if we found a subject objectively demonstrated clear low-effort throughout the experiment, i.e., failed all attention checks, we rejected the submission. The rejected submissions will be removed immediately, and Prolific will automatically recruit another subject as substitution. Therefore, we only paid for approved submissions and no further data screening procedure needed to be applied on the collected data.

In overall, there were a total of 3200 lineup evaluations made by 160 subjects in both experiment I and experiment II respectively, where 320 lineup evaluations were attention checks and were not used in the following analysis.

The collated dataset is provided in `vi_survey` of the `visage` `R` package.

@@

Figure \ref{fig:power-poly-uniform} shows the estimated power of visual test against natural logarithm of the effect with comparison to the power of an exact test - F-test, and the power of two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions. In overall, the power of all four tests increases as the effect becomes larger. The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while others respond inactively to the change of effect and remain lower than 25% throughout the period, showing that as an exact test, the F-test is relatively more sensitive to the type of model defects that being considered. The power of visual test arises steadily and nearly linearly to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features. Other two inappropriate conventional tests shows improvement at the same time but at a lower rate. This coincides the point made by @cook1982residuals mentioned in \ref{hypothesis-testing} that residual-based tests for a specific type of model defect are sensitive to other types of model defects. At $log_e(E) = 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin. The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 30% respectively.

What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test. The margin is largest at around $log_e(E) = 2$. An example lineup is included in Figure \ref{fig:power-overview} where none of subjects detect the actual data plot positioned at panel 14. It demonstrates that at this level of difficulty, the designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model. From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public even though it is true. At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger. At this level of difficulty, the designed visual features are usually detectable but it may not stand out from the lineup as other null plots may happen to include outliers or visual patterns that are are considered to be more attractive by human, and thus recognized as the most different plot. Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison. The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 20, where two out of five subjects detect it. It can be observed that a M-shape is presented in plot 20, but the signal is not strong enough to attract all five subjects, resulting in a visual p-value sightly above the desired significance level $\alpha = 0.05$. At $log_e(E) = 4$ and $log_e(E) = 6$, the designed visual features become much clear and attractive, leading to a high percentage of rejection of the null hypothesis. Figure \ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It is also interesting to observe that the power curve of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible.

@@

```{r fig:power-poly-uniform, fig.cap = "..."}
vi_survey_processed %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) + 
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                pivot_longer(F_p_value:SW_p_value), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  theme_light(base_size = 5) -> p

VI_MODEL$plot_lineup(vi_lineup[["poly_39"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(base_size = 5),
                     remove_axis = TRUE)

VI_MODEL$plot_lineup(vi_lineup[["poly_415"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(base_size = 5),
                     remove_axis = TRUE) -> p1

VI_MODEL$plot_lineup(vi_lineup[["poly_131"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(base_size = 5),
                     remove_axis = TRUE) -> p2

VI_MODEL$plot_lineup(vi_lineup[["poly_322"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(base_size = 5),
                     remove_axis = TRUE) -> p3

VI_MODEL$plot_lineup(vi_lineup[["poly_378"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(base_size = 5),
                     remove_axis = TRUE) -> p4

design <- c(patchwork::area(1, 1, 1, 8),
            patchwork::area(2, 1, 2, 4), 
            patchwork::area(2, 5, 2, 8), 
            patchwork::area(3, 1, 3, 4), 
            patchwork::area(3, 5, 3, 8))

patchwork::wrap_plots(p, B = p1, C = p2, D = p3, E = p4, design = design)
```

## Distribution of regressor

The impact of the distribution of $X_raw$ on the power is shown in Figure \ref{fig:dist-power}. The power curve of F-test is stable across different distributions, while the visual test has a steeper power curve for normal and uniform distribution. BP-test performs worse for discrete uniform distribution and uniform distribution but has relatively high power for normal distribution. SW-test outperforms BP-test for discrete uniform distribution but remains as the worst test for other distributions. The results indicate those inappropriate residual-based tests are sensitive to the distribution of the regressor.

```{r dist-power, fig.cap="..."}
vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) +
  stat_smooth(geom = "line",
              data = ~boot_y(.x, reject_dirichlet, times = 500),
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"),
              method = "glm",
              method.args = list(family = binomial),
              se = FALSE,
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                pivot_longer(F_p_value:SW_p_value),
              aes(log_effect_size, as.numeric(value < 0.05), col = name),
              method = "glm",
              method.args = list(family = binomial),
              se = FALSE) +
  facet_wrap(~x_dist) +
  theme_light(base_size = 5)
```

@@
